Categorical Predictor Embeddings and Attention Based Neural Networks in Actuarial Modeling

### 2021-04-07 manuscript/manuscript.Rmd
Add Section 6 and 7

### Model 6: TabTransformer
Whereas Model 5 makes a simple modification to the structure of Model 4, the TabTransformer model is slightly more complex. With respect to
similarities between Model 5 and 6, both use embeddings with a standard 8 dimensions for each categorical covariate. In terms of differences,
Model 6 uses a Transformer layer in place of self-attention and a positional encoding of a single dimension, i.e. $d_col = 1$. The TabTransformer
architecture is shown in Figure Figure \@ref(fig:nn-3).

@@

## Training and Inference {#training-inference-attn}
We follow a similar training setup for the three attention based models, to the neural networks discussed in the previous section. For Models 5 and 6, 
we utilize the \texttt{Adam} optimizer as before, however, due to the increased complexity of the models in this section,
we use learning rate of 0.001 (compared to 0.01 for the simpler models). Reflecting this reduced learning rate, 
we train on the analysis set for a maximum of 150 epochs, and select the best model based on the lowest loss achieved on the assessment. Moreover, since these models are
more complex than those tested in the previous section, we also add some minimal regularization using dropout [@srivastava2014dropout] to the attention layers 
and the last layer of the network. Dropout functions by setting some components of the network randomly to zero during each epoch of training. This forces each
component of the network to learn an independent representation of the data. In this section, we drop out components with a probability of 2.5\%. A final detail is that, 
for both of these models, we use an embedding dimension of size 16 for each of the categorical covariates, since the attention mechanism requires that the embeddings 
be of a common length. For the TabTransformer model, we use a row embedding of dimension 4.

@@

In Table \@ref(tab:cv-results-attn), we show the cross-validated performance metrics of each of the models, 
as well as the two best performing models from the previous section. As before, we report both the root mean squared error (RMSE) 
and the mean absolute error (MAE). The results show that Model 5, which simply adds attention to Model 4, achieves slightly worse
performance than Model 4 as measured by the RMSE, but using the MAE as a metric, outperforms Model 4. Thus, there is some room to argue that
for tabular data modelling that uses embeddings, the use of attention within models should be considered. Disappointingly, the TabNet 
architecture performs poorly, perhaps due to insufficient hyperparameter tuning, but indicating that ``out of the box", TabNet does not
appear to work well for this modelling task. Finally, the TabTransformer model shows the best performance of all models considered in 
this research, beating the others by a significant margin on both the RMSE and MAE metrics.

@@

|Model                                                                          |   RMSE|    MAE|
|:-----------------------------------------------------------------------------|------:|------:|
|Model 4: MLP with multidimensional embeddings                                 | 60,973| 37,574|
|Model 5: Simple Attention                                                     | 61,022| 37,025|
|Model 6: TabNet                                                               | 62,938| 38,386|
|Model 7: TabTransformer                                                       | 59,900| 36,343|
|Linear regression, predictions capped below at 0.01                           | 60,879| 38,431|

@@

## Explaining Attention Based models

@@

Conclusions

>>>>

### Model 6: TabTransformer
Whereas Model 5 makes a simple modification to the structure of Model 4, the TabTransformer model is slightly more complex. With respect to
similarities between Model 5 and 6, both use embeddings with a standard 8 dimensions for each categorical covariate. In terms of differences,
Model 6 uses a Transformer layer in place of self-attention and a positional encoding of a single dimension, i.e. $d_col = 1$. The TabTransformer
architecture is shown in Figure \@ref(fig:nn-3).

@@

## Training and Inference {#training-inference-attn}
We follow a similar training setup for the three attention based models, to the neural networks discussed in the previous section. For Models 5 and 6, 
we utilize the \texttt{Adam} optimizer as before, however, due to the increased complexity of the models in this section,
we use learning rate of 0.001 (compared to 0.01 for the simpler models). Reflecting this reduced learning rate, 
we train on the analysis set for a maximum of 150 epochs, and select the best model based on the lowest loss achieved on the assessment. Moreover, since these models are
significantly more complex than those tested in the previous section, we also add some minimal regularization using dropout [@srivastava2014dropout] to the attention layers 
and the last layer of the network, to prevent overfitting. Dropout functions by setting some components of the network randomly to zero during each epoch of training. 
This forces each component of the network to learn a somewhat independent representation of the data, thus helping to avoid overfitting by preventing components of the 
network from becoming too specialized. In this section, we drop out components with a probability of 2.5\%. A final detail is that, 
for both of these models, we use an embedding dimension of size 16 for each of the categorical covariates, since the attention mechanism requires that the embeddings 
be of a common length. For the TabTransformer model, we use a row embedding of dimension 4.

@@

In Table \@ref(tab:cv-results-attn), we show the cross-validated performance metrics of each of the models, 
as well as the two best performing models from the previous section. As before, we report both the root mean squared error (RMSE) 
and the mean absolute error (MAE). The results show that Model 5, which simply adds attention to Model 4, outperforms Model 4 on both the RMSE and 
MAE metrics. Thus, there is some room to argue that for tabular data modelling that uses embeddings for categorical data, 
the use of attention within models should be considered. Disappointingly, the TabNet architecture performs poorly, perhaps due to insufficient hyperparameter tuning, 
but indicating that ``out of the box", TabNet does not appear to work well for this modelling task. Finally, the TabTransformer model shows the best performance of all
 models considered in this research, beating the other models by a significant margin on both the RMSE and MAE metrics.

@@

|Model                                                                          |   RMSE|    MAE|
|:-----------------------------------------------------------------------------|------:|------:|
|Model 4: MLP with multidimensional embeddings                                 | 60,973| 37,574|
|Model 5: Simple Attention                                                     | 60,601| 36,739|
|Model 6: TabNet                                                               | 62,938| 38,386|
|Model 7: TabTransformer                                                       | 59,900| 36,343|
|Linear regression, predictions capped below at 0.01                           | 60,879| 38,431|

@@

## Exploring the TabTransformer model

Here, we explore what the TabTransformer model has learned from the data via the attention process that forms the core of the model. We focus on the difference 
between the flood embedding learned by the model, compared with the output of the Transformer layer that processes the embeddings before these enter the model. 
@Huang2020 call the latter ``contextual embeddings", because the attention mechanism within the TabTransformer model augments the embeddings based on the 
other covariates that appear for each observation $i$.

To illustrate this, we selected a single TabTransformer fit on one fold of the training data, and selected at random 10 000 individual observations. For each
observation, the embedding relating to the flood zone of the observation was extracted from the model, as well as the contextual embedding that was produced 
for that observation by the TabTransformer model. We found that a PCA and t-SNE analysis of the embeddings produced very similar results to those shown in the previous section,
so we do not show these plots here. To investigate the difference between the embeddings and contextual embeddings for the `flood_zone` variable, 
we show the first component from a PCA analysis of each of these in Figure \@ref(fig:embd).

```{r embd, echo = FALSE, fig.cap = "Embedding versus Contextual Embedding, Flood Zone, first PCA components", out.width = "80%", fig.align = "center"}
knitr::include_graphics("images/emb_vs_cembd.pdf")
```

The figure shows that, whereas the first principal component of the embeddings occupy discrete points only (shown on the $x$-axis), 
the first principal component of the **contextual** embeddings is continuous (shown on the $y$-axis), meaning that the model has modified the values
of the embedding in a smooth manner. Since the inputs to the attention mechanism which is responsible for this are the embeddings for the other categorical
variables for each observation, these other variables have influenced the final embedding used within the model for the `flood_zone` variable. By considering 
each of the other categorical variables entering the model, it was found that the basement enclosure crawl space variable explains most of the variation
observed within the flood zone embeddings, as shown in Figure \@ref(fig:embdexp). The attention mechanism produces the lowest values for those observations
with an unfinished basement, whereas the model produces the highest values for buildings without a crawl space.

```{r embdexp, echo = FALSE, fig.cap = "Embedding versus Contextual Embedding, Flood Zone, first PCA components, points coloured based on crawl space variable", out.width = "80%", fig.align = "center"}
knitr::include_graphics("images/emb_vs_cembd_exp.pdf")
```
To investigate the predictive value added by the contextual embeddings, a simple Gamma GLM was fit to these 10 000 observations, with the goal of predicting the 
claim amount based only on either the embeddings or the contextual embeddings. As shown in Table \@ref(tab:emd-vs-cembd), the performance of the latter model, 
as indicated by the Aikaike Information Criterion (AIC) was significantly better than the latter model (37 129 compared with 37 463), 
in other words, the contextual embeddings that incorporate information across all of the categorical variables are significantly better predictors 
of claim severity compared with the embeddings.

Table: (\#tab:emd-vs-cembd) AIC for GLMs predicting claim severity, embeddings compared with contextual embeddings.

|Embedding Type       |      AIC|
|:--------------------|--------:|
|Embedding            | 37,464|
|Contextual embedding | 37,129|

@@

This study has focused on how embeddings can be applied within the context of actuarial modelling, by illustrating a range of approaches: first, relatively simple
neural networks with embeddings were defined, the embeddings from these models were reused within traditional GLM models and, finally, more advanced attention based models 
were studied. These models were demonstrated within the context of a claims severity modelling problem based on data from the NFIP. Whereas we have only considered severity 
in this study, it is likely that our findings will replicate for the modelling of claim frequency, which is usually an easier problem. Our key findings are that 
processing categorical variables with multi-dimensional embeddings leads to enhanced performance within simple neural networks, and that, by including these within traditional
GLM models, the predictive performance of these models can be enhanced, compared to the usual encoding scheme for categorical variables. Some of the attention based models 
considered within this study performed well, in particular, the TabTransformer model of @Huang2020 appears quite promising for use in actuarial modelling. The study has also 
illustrated how embeddings can be visualized using the PCA and t-SNE techniques. We found that, in the case of the NFIP data, it is possible to explain which variables modify 
the embeddings based on the context of each observation.

Several avenues for future research can be considered. One surprising finding of our work is that the TabNet model performed quite poorly, leading us to conclude that this 
model appears to be less suitable for actuarial modelling. Understanding these findings in more detail would be a valuable contribution to the literature. Another helpful contribution
would be to investigate the optimal hyperparameters for the TabTransformer model in the context of actuarial modelling. Finally, extneding our findings to a claims frequency example
could be considered.
